---
title: "Production"
author: "gabriel"
date: "2/8/2019"
output:
  rmdformats::readthedown:
    thumbnails: true
    lightbox: true
    toc_depth: 3
    gallery: true
    highlight: tango

---

## loading packages and data

```{r, warning=F, message=F}
pacman:: p_load(readr, h2o, dplyr,ggplot2, caret, imputTS, randomForest, data.table, lubridate, corrplot, quantreg, iClick, h2o)

features_tr <- read_csv("./input/dengue_features_train.csv", 
    col_types = cols(week_start_date = col_date(format = "%Y-%m-%d")))

labels_tr <- read_csv("./input/dengue_labels_train.csv")

features_te <- read_csv("./input/dengue_features_test.csv", 
    col_types = cols(week_start_date = col_date(format = "%Y-%m-%d")))

submission_format <- read_csv("./input/submission_format.csv")


options(scipen=999)



```

create two dataset, train and test

```{r, warning=F}

train <- left_join(x = features_tr, y = labels_tr, by = c("year", "weekofyear", "city"))

test <- left_join(x = features_te, y = submission_format, by = c("year", "weekofyear", "city"))

train$type <- "train"

test$type <- "test"

all <- rbind(train, test)

all %>% group_by(city,year, week_start_date) %>%  mutate(avg_temp = mean(reanalysis_avg_temp_k, reanalysis_air_temp_k, na.rm = T, trim = 0), avg_precip = mean(station_precip_mm, precipitation_amt_mm, na.rm = T, trim = 0),  station_avg_temp_c = NULL, reanalysis_avg_temp_k = NULL, reanalysis_air_temp_k = NULL, station_precip_mm = NULL, precipitation_amt_mm = NULL) # Create avg of tempearature and preciptation


```


```{r, warning=F}

train$type <- "train"

test$type <- "test"

all <- rbind(train, test)

```

## Split data by city 

```{r, warning=F}

sj <-  all %>% filter(city == "sj" | type == "train") 

```


The missing values will be replaced by a forecast

```{r}

sj <- cbind(sj[ , c("city", "type", "week_start_date")] ,  apply(X = sj[, which(sapply(sj,is.numeric))], MARGIN = 2, function(x) imputeTS::na.kalman(x)))

```


Iquitos

```{r}

iq <- all %>% filter(city == "iq" | type == "train")

iq <- cbind(iq[ , c("city", "type", "week_start_date")] ,  apply(X = iq[, which(sapply(iq,is.numeric))], MARGIN = 2, function(x) imputeTS::na.kalman(x)))

```


# Modelling San Juan


```{r, warning=F}

sj$city <- NULL


sj$weekofyear <- as.character(sj$weekofyear) 

```

train model to predict Test (All train)

```{r, warning=F, message=F, echo=T}


# Get the best mtry train a random forest using that mtry

sj$year <- as.character(sj$year)

# preprocessparams <- preProcess(x = select(sj[ , which(sapply(sj,is.numeric))], -total_cases) ,method = c("center", "scale"))
# 
# sj <- predict(preprocessparams, sj)

sj$type <- NULL
#sj$week_start_date <- NULL


```

```{r}

h2o.init()

#data to h2o cluster
train.h2o <- as.h2o(sj)

x.indep <- which(names(train.h2o) != "total_cases")

y.dep <- which(names(train.h2o) == "total_cases")  

deep_learn <- h2o.glm(y=y.dep, x=x.indep, training_frame = train.h2o)

h2o.performance(deep_learn)


```


### Make final prediction for San Juan (test)

```{r, warning=F}

sj_te <- all %>% filter(city == "sj" & type == "test")

sj_te$year <- as.character(sj_te$year)


```

The missing values will be replaced by a forecast

```{r, warning=F}

sj_te <- cbind(sj_te[ , c("city", "type", "week_start_date","year")] ,  apply(X = sj_te[, which(sapply(sj_te,is.numeric))], MARGIN = 2, function(x) imputeTS::na.kalman(x)))

```

### Apply parameters and predict

```{r, warning=F}


test.h2o <- as.h2o(sj_te)

predict.deep_learn <- as.data.frame(h2o.predict(deep_learn, test.h2o))

sj_te$total_cases <- predict.deep_learn$predict

```

# Modelling Iquito

Iquito

```{r}

iq$weekofyear <- as.character(iq$weekofyear) 

```


train model to predict Test (All train)

```{r, warning=F, echo=F}


iq$city <- NULL

iq$type <- NULL

iq$year <- as.character(iq$year)


#data to h2o cluster
train.h2o <- as.h2o(iq)

x.indep <- which(names(train.h2o) != "total_cases")

y.dep <- which(names(train.h2o) == "total_cases")  

deep_learn <- h2o.glm(y=y.dep, x=x.indep, training_frame = train.h2o)

h2o.performance(deep_learn)

```


### Make final prediction for Iquito (test)

```{r}

iq_te <- all %>% filter(city == "iq" & type == "test")

iq_te$weekofyear <- as.character(iq_te$weekofyear)

```


The missing values will be replaced by a forecast

```{r, warning=F}

iq_te <- cbind(iq_te[ , c("city", "type", "week_start_date", "weekofyear")] ,  apply(X = iq_te[, which(sapply(iq_te,is.numeric))], MARGIN = 2, function(x) imputeTS::na.kalman(x)))


```


```{r}
iq_te$type <- NULL
iq_te$city <- NULL
#iq_te$year <- NULL
#iq_te$week_start_date <- NULL




```

Predict iq

```{r, warning=F}



test.h2o <- as.h2o(iq_te)

predict.deep_learn <- as.data.frame(h2o.predict(deep_learn, test.h2o))

iq_te$total_cases <- predict.deep_learn$predict



```

# Final Submission using RF

```{r, warning=F}


iq_te$city <- "iq" 

iq_te$type <- "test"

rf_final_df <- rbind(iq_te[ , c("city", "type", "total_cases","weekofyear", "week_start_date", "year")], 
                     sj_te[ , c("city", "type", "total_cases","weekofyear", "week_start_date", "year")])

submission_format$total_cases <- 10


submission_format$year <- as.character(submission_format$year)

submission_format$weekofyear <- as.character(submission_format$weekofyear)

submission_rf <- full_join(x = submission_format, y = rf_final_df, by = c("year","weekofyear", "city"))

data.table::setnames(submission_rf, old = "total_cases.y", "total_cases")

submission_rf$total_cases <- base::round(submission_rf$total_cases,0) # Round


submission_rf$total_cases[submission_rf$total_cases < 0 ] <- 0

#base::which(submission_rf$total_cases < 0) <- 0
#if negative, change to 0
```


```{r}

ggplot(sj, aes(week_start_date, y = total_cases)) + geom_line()
ggplot(iq, aes(week_start_date, y = total_cases)) + geom_line()

ggplot(submission_rf, aes(week_start_date, total_cases)) + geom_line() + facet_wrap(city~.)

submission_rf <- submission_rf %>% select(city, year, weekofyear, total_cases) 

write.csv(submission_rf, "submission_rf.csv", quote = F, row.names = F)

summary(submission_format)

summary(submission_rf)


```

```{r}
```

